---
title: "Activity - Project 1"
---

This markdown document demonstrates the steps necessary to complete Project #1 in the Coursera Reproducible Research Course

First we will load required packages
```{r}
library(knitr)
library(dplyr)
library(ggplot2)
```
Then we will load and look at a sample of the data

```{r}
data<-read.csv("activity.csv")
head(data)
```

Next we will remove the NA values from the steps column in the data 

```{r}
clean_data <- data[ with (data, { !(is.na(steps)) } ), ]
```

Next lets take a look at the data to ensure NA values were removed properly

```{r}
head(clean_data)
```

**Now we can perform some analysis of the data**


*What is the Mean number of steps per day?*

Sum total steps per day and show sample

```{r}
daydata <- group_by(clean_data, date)
totalsteps<- summarize(daydata, total = sum(steps))
head(totalsteps)
```

Histogram of total steps per day

```{r}
hist(totalsteps$total, main="Total steps per day", 
     xlab="Steps")
```


Summary of Total Steps per day
```{r}
summary(totalsteps)
```

*What is the average daily activity?*

Prepare data

```{r}
intervaldata <- aggregate(steps ~ interval, data, mean)
```

Make a time series plot

```{r}
plot(intervaldata$interval, intervaldata$steps, type='l', 
     main="Average steps for all days", xlab="Interval", 
     ylab="Average steps")
```

Find Maximum

```{r}
maxsteps <- which.max(intervaldata$steps)
intervaldata[maxsteps, ]
```

*Impute Missing Values*

Calcualte total missing

```{r}
sum(is.na(data))
```

Replace NA values with 5 minute interval average, show sample of resulting data

```{r}
No_NA <- data
for (i in 1:nrow(No_NA)) {
  if (is.na(No_NA$steps[i])) {
    interval_value <- No_NA$interval[i]
    steps_value <- intervaldata[
      intervaldata$interval == interval_value,]
    No_NA$steps[i] <- steps_value$steps
  }
}

head(No_NA)
```

calculate total steps and show sample

```{r}
NO_NA_Steps<- aggregate(steps ~ date, No_NA, sum)
head(NO_NA_Steps)

```

Make histogram

```{r}
hist(NO_NA_Steps$steps, main="Total steps per day", 
     xlab="Steps")

```

Show mean of imputed data
```{r}
mean(NO_NA_Steps$steps)
```

Show median of imputed data
```{r}
median(NO_NA_Steps$steps)
```

Show mean of data with NA removed
```{r}
mean(totalsteps$total)
```

Show median of  data with NA removed
```{r}
median(totalsteps$total)
```

Median values are slightly different and mean values are exactly the same

*Are there different patterns between weekends and week days?*

Prepare data by categorizing days
```{r}
No_NA['type_of_day'] <- weekdays(as.Date(No_NA$date))
No_NA$type_of_day[No_NA$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
No_NA$type_of_day[No_NA$type_of_day != "weekend"] <- "weekday"
```


convert to factor and calculate average steps
```{r}
No_NA$type_of_day <- as.factor(No_NA$type_of_day)

Steps_by_interval <- aggregate(steps ~ interval + type_of_day, No_NA, mean)

head(Steps_by_interval)
```

Plot data on weekdays vs weekends

```{r}
qplot(interval, steps, data = Steps_by_interval,geom=c("line"), xlab = "Interval",  ylab = "Steps", main = "") +
  facet_wrap(~ type_of_day, ncol = 1)
```
